
======================================================================

# 2024 Institute for Clinical Evaluative Sciences. All rights reserved.

# TERMS OF USE:

##Not for distribution.## This code and data is provided to the user solely for its own non-commerical use by individuals and/or not-for-profit corporations. Users shall not distribute without express written permission from the Institute for Clinical Evaluative Sciences.

##Not-for-profit.## This code and data may not be used in connection with profit generating activities.

##No liability.## The Institute for Clinical Evaluative Sciences makes no warranty or representation regarding the fitness, quality or reliability of this code and data.

##No Support.## The Institute for Clinical Evaluative Sciences will not provide any technological, educational or informational support in connection with the use of this code and data.

##Warning.## By receiving this code and data, user accepts these terms, and uses the code and data, solely at its own risk.

======================================================================





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pacman)
p_load("dplyr", "tidyverse", "data.table", "base", "utils", "table1", "tableone", "ggpubr", "ggplot2", "tidyr", "survival", "survminer", "flexsurv", "flexsurvcure", "mixcure", "lubridate", "plyr", "Epi", "mgcv", "splines", "matrixStats", "stringr", "forecast", "scales", "tsibble", "ggforce")

```

#import StatCan population data

```{r}
#StatCanpop <- read.csv("REDACTED")

popON_M <- filter(StatCanpop, GEO == "Ontario", Sex == "Males")
popON_M <- filter(popON_M, !Age_group %in% c('0 to 4 years', '5 to 9 years', '10 to 14 years', '0 to 14 years'))
popON_M <- popON_M[c(-2, -3)]
popON_M$A <- as.numeric(rep(0:14, each = 50, times = 1))
popON_M <- popON_M[c(-2)]
popON_M <- popON_M %>% dplyr::rename(P = REF_DATE, N_July1 = VALUE)

popON_M$N[1] <- popON_M$N_July1[1]
for (i in 1:nrow(popON_M)) {
  
  if(popON_M$P[i] ==  1971) {
    popON_M$N[i] <- popON_M$N_July1[i]
  } else{
    popON_M$N[i] <- (popON_M$N_July1[i-1] + popON_M$N_July1[i])/2
  }
}
popON_M <- popON_M[c(-2)]
popON_M$N <- round(popON_M$N)
popON_M <- popON_M %>% dplyr::rename(Y = N)

#####
popON_F <- filter(StatCanpop, GEO == "Ontario", Sex == "Females")
popON_F <- filter(popON_F, !Age_group %in% c('0 to 4 years', '5 to 9 years', '10 to 14 years', '0 to 14 years'))
popON_F <- popON_F[c(-2, -3)]
popON_F$A <- as.numeric(rep(0:14, each = 50, times = 1))
popON_F <- popON_F[c(-2)]
popON_F <- popON_F %>% dplyr::rename(P = REF_DATE, N_July1 = VALUE)

popON_F$N[1] <- popON_F$N_July1[1]
for (i in 1:nrow(popON_F)) {
  
  if(popON_F$P[i] ==  1971) {
    popON_F$N[i] <- popON_F$N_July1[i]
  } else{
    popON_F$N[i] <- (popON_F$N_July1[i-1] + popON_F$N_July1[i])/2
  }
}
popON_F <- popON_F[c(-2)]
popON_F$N <- round(popON_F$N)
popON_F <- popON_F %>% dplyr::rename(Y = N)

### BOTH SEXES
popON <- filter(StatCanpop, GEO == "Ontario", Sex == "Both sexes")
popON <- filter(popON, !Age_group %in% c('0 to 4 years', '5 to 9 years', '10 to 14 years', '0 to 14 years'))
popON <- popON[c(-2, -3)]
popON$A <- as.numeric(rep(0:14, each = 50, times = 1))
popON <- popON[c(-2)]
popON <- popON %>% dplyr::rename(P = REF_DATE, N_July1 = VALUE)

popON$N[1] <- popON$N_July1[1]
for (i in 1:nrow(popON)) {
  
  if(popON$P[i] ==  1971) {
    popON$N[i] <- popON$N_July1[i]
  } else{
    popON$N[i] <- (popON$N_July1[i-1] + popON$N_July1[i])/2
  }
}
popON <- popON[c(-2)]
popON$N <- round(popON$N)
popON <- popON %>% dplyr::rename(Y = N)
```

#create risk time ("Person-Years") in Lexis triangles from population count data.
```{r}
childpop_M_Y <- N2Y(data = popON_M, return.dfr = TRUE)

childpop_F_Y <- N2Y(data = popON_F, return.dfr = TRUE)

childpop_Y <- N2Y(data = popON, return.dfr = TRUE)

```

#Load individual-level patient data with incidence characteristics. 
#We utilized incident cases of primary childhood cancer from 1985-2019
#Data source - Pediatric Oncology Group of Ontario Networked Information System (POGONIS)

```{r}

incidence_IPD <- read.csv("REDACTED")
incidence <- as.data.frame(incidence_IPD)

incidence <- filter(incidence, dxyear > 1984) #GAM models are informed by individual-level data for patients diagnosed from 1985-2019

source('GAM_Functions.R')

```

#create new cancer type grouping variable for incidence analyses
```{r}

incidence <- incidence %>% mutate(ICCC_regroup = case_when(
    iccc_subgroup_code %in% c("Ia.") ~ "ALL",
    iccc_subgroup_code %in% c("Ib.", "Ic.", "Id.", "Ie.") ~ "AML and other leukemias",
    iccc_subgroup_code %in% c("IIa.") ~ "Hodgkin lymphomas",
    iccc_subgroup_code %in% c("IIb.", "IIc.", "IId.", "IIe.") ~ "Non-Hodgkin lymphomas and other lymphomas",
    iccc_subgroup_code %in% c("IIIb.") ~ "Astrocytoma",
    iccc_subgroup_code %in% c("IIIa.", "IIIc.", "IIId.", "IIIe.", "IIIf.") ~ "Other CNS neoplasms",
    iccc_subgroup_code %in% c("IVa.", "IVb.") ~ "Neuroblastoma",
    iccc_subgroup_code %in% c("V.") ~ "Retinoblastoma",
    iccc_subgroup_code %in% c("VIa.", "VIb.", "VIc.") ~ "Renal tumours",
    iccc_subgroup_code %in% c("VIIa.", "VIIb.", "VIIc.") ~ "Hepatic tumours",
    iccc_subgroup_code %in% c("VIIIa.", "VIIIb.", "VIIIc.", "VIIId.", "VIIIe.") ~ "Bone tumours",
    iccc_subgroup_code %in% c("IXa.", "IXb.", "IXc.", "IXd.", "IXe.") ~ "Soft tissue sarcomas",
    iccc_subgroup_code %in% c("Xa.", "Xb.", "Xc.", "Xd.", "Xe.") ~ "Germ cell tumours",
    iccc_subgroup_code %in% c("XIa.", "XIb.", "XIc.", "XId.", "XIe.", "XIf.", "XIIa.", "XIIb.") ~ "Other epithelial and unspecified neoplasms"))

incidence$ICCC_regroup <- factor(incidence$ICCC_regroup,
                               levels = c("ALL", "AML and other leukemias", "Hodgkin lymphomas", "Non-Hodgkin lymphomas and other lymphomas", "Astrocytoma", "Other CNS neoplasms", "Neuroblastoma", "Retinoblastoma", "Renal tumours", "Hepatic tumours", "Bone tumours", "Soft tissue sarcomas", "Germ cell tumours", "Other epithelial and unspecified neoplasms"),
                               labels = c("ALL", "AML and other leukemias", "Hodgkin lymphomas", "Non-Hodgkin lymphomas and other lymphomas", "Astrocytoma", "Other CNS neoplasms", "Neuroblastoma", "Retinoblastoma", "Renal tumours", "Hepatic tumours", "Bone tumours", "Soft tissue sarcomas", "Germ cell tumours", "Other epithelial and unspecified neoplasms"))


```

#Incidence counts
```{r}
################ generating list of all possible combinations ##########
####### Sex (2), dxyear (35), age (15), iccc regroup (14)
####### Incidence by age, sex, dxyear, ICCC group between 1985-2019 

count_NA <- expand.grid(dxyear = unique(incidence$dxyear),
                        sex = unique(incidence$sex),
                        age = unique(incidence$age),
                        ICCC_regroup = unique(incidence$ICCC_regroup),stringsAsFactors = F) %>% as.data.frame()

count <- incidence %>%
  group_by(dxyear, sex, age, ICCC_regroup) %>%
  dplyr::summarise(N = n()) %>% ungroup()

new_count <- count_NA %>%
  left_join(count)

new_count$N[is.na(new_count$N)] <- 0

```


#expanded grid count data
```{r}
firstyr <- 1985
lastyr <- 2019
ageyrs <- ((lastyr - firstyr)+1)*2

A <- as.numeric(rep(0:14, each = 70))
lexisdata <- data.frame(A)
lexisdata$P <- as.numeric(rep(firstyr:lastyr, each = 2, times = 15))
lexisdata$C <- rep(NA)
lexisdata$up <- as.numeric(rep(c(0:1)))
lexisdata$C <- as.numeric((lexisdata$P - lexisdata$A - lexisdata$up))

```

#create A , P , C  vars
```{r}

incidence$A <- incidence$age
incidence$P <- incidence$dxyear
incidence$C <- incidence$birth_year

```


#create and store subsetted dataframes
```{r}
inc <- data.frame(incidence)
tomodel <- inc %>% select(ikn, patient_id, sex, age, agegroups, agegroup, dxyear, birth_year, A, P, C, iccc_group_code, iccc_subgroup_code, ICCC_regroup)

tomodel1 <- tomodel %>% dplyr::count(sex, ICCC_regroup)
```

#run list
```{r}
res.list <- vector(mode = "list")
for(i in 1:dim(tomodel1)[1]){
  
df_tosave <- filter(incidence, sex == tomodel1$sex[i], ICCC_regroup == tomodel1$ICCC_regroup[i]) %>% group_by(P, C, A) %>% dplyr::summarise(N = n())

df_tosave$up <- df_tosave$P - df_tosave$A - df_tosave$C
df_tosave <- lexisdata %>% left_join(df_tosave)
df_tosave[is.na(df_tosave)] <- 0
df_tosave <- transform(df_tosave, A = A+(1+up)/3,
                            P = P+(2-up)/3,
                            C = C+(1+up)/3)
  
if(tomodel1$sex[i] == "Male"){
  df_tosave <- left_join(df_tosave, childpop_M_Y, by=c("A", "P"))
}
else if (tomodel1$sex[i] == "Female") {
df_tosave <- left_join(df_tosave, childpop_F_Y, by=c("A", "P"))
}

res.list[[i]] <- df_tosave

names(res.list)[i] <- paste0(tomodel1$ICCC_regroup[i], "_", tomodel1$sex[i])

res.list[[i]] <- as.data.frame(res.list[[i]])

}

```

#######################################################################################
#### Cross-validation #####
#######################################################################################

#1. split the dataset into multiple training sets (by id), save in a list
```{r}
trainingsets <- vector(mode = "list")

for(i in 1:length(res.list)){

  data_tr <- data_tr_f(res.list[[i]]) #stretch the dataframes that are needed to then create the 21 split dataframes for each

  trainingsets[[i]] <-  list(data_tr = data_tr)
}
names(trainingsets) <- names(res.list)

#saveRDS(trainingsets, file = "trainingsets.Rds")


```

#2. create list within a list: for every data_tr dataframe in trainingsets, split it up and save 21 new dataframes under each of the 24 named cancer type/sex combinations
```{r}
split_trainingsets <- vector(mode = "list")

for(i in 1:length(trainingsets)){

splits <- vector(mode = "list")
for(j in 1:dim(ids)[1]){

  df_byid  <- trainingsets[[i]][["data_tr"]] %>% filter(.id == ids$ids[j])
  
  splits[[j]] <- as.data.frame(df_byid)
}
  split_trainingsets[[i]] <- splits
}
names(split_trainingsets) <- names(trainingsets)

#saveRDS(split_trainingsets, file = "split_trainingsets.Rds")


```


#3. run each of the 21 training datasets (saved in split_trainingsets list) through the gam model(s), prediction functions in order to calculate RMSE at the end. 
# Save the RMSE values in a list according to gam model fit so that gam model with lowest RMSE can be run in step 4.
```{r}
options(warn = 0)

newlist_tscv5 <- vector(mode = "list")

for(i in 1:length(split_trainingsets)){ #length = 28 cancertype/sex groups

results_tscv5 <- vector(mode = "list")

for(j in 1:length(split_trainingsets[[i]])){ #run the 3 functions below on EACH of the 21 datasets under each of the 28 cancertype/sex combinations

rates_tscv <- rates_tscv_f(split_trainingsets[[i]][[j]]) 
Pfr_tscv5 <- Pfr_tscv5_f(split_trainingsets[[i]][[j]]) #predict 5th year ahead
df_models_tscv5 <- gammodels_tscv5_f(split_trainingsets[[i]][[j]]) #predict 5th year ahead

results_tscv5[[j]] <-  list(rates_tscv = rates_tscv, df_models_tscv5 = df_models_tscv5)

}
newlist_tscv5[[i]] <-  results_tscv5
}
names(newlist_tscv5) <- names(split_trainingsets)

#summary(warnings(newlist_tscv5))
```

#4. RMSE results for 5-year ahead TSCV (time-series cross-validation)
```{r}
RMSEresults_tscv5 <- vector(mode = "list")

for(i in 1:length(newlist_tscv5)){ #length = 28 cancertype/sex groups

RMSEs_tscv5 <- vector(mode = "list") #create a new list that contains the RMSE results for each of the 28 cancertype/sex groups

df_models1_bind <- newlist_tscv5[[i]][[1]][["df_models_tscv5"]][[1]]
df_models2_bind <- newlist_tscv5[[i]][[1]][["df_models_tscv5"]][[2]]
df_models3_bind <- newlist_tscv5[[i]][[1]][["df_models_tscv5"]][[3]]
df_models4_bind <- newlist_tscv5[[i]][[1]][["df_models_tscv5"]][[4]]
df_models5_bind <- newlist_tscv5[[i]][[1]][["df_models_tscv5"]][[5]]
df_models6_bind <- newlist_tscv5[[i]][[1]][["df_models_tscv5"]][[6]]

for(k in 2:21){  
df_models1_bind <- rbind(df_models1_bind, newlist_tscv5[[i]][[k]][["df_models_tscv5"]][[1]])
df_models2_bind <- rbind(df_models2_bind, newlist_tscv5[[i]][[k]][["df_models_tscv5"]][[2]])
df_models3_bind <- rbind(df_models3_bind, newlist_tscv5[[i]][[k]][["df_models_tscv5"]][[3]])
df_models4_bind <- rbind(df_models4_bind, newlist_tscv5[[i]][[k]][["df_models_tscv5"]][[4]])
df_models5_bind <- rbind(df_models5_bind, newlist_tscv5[[i]][[k]][["df_models_tscv5"]][[5]])
df_models6_bind <- rbind(df_models6_bind, newlist_tscv5[[i]][[k]][["df_models_tscv5"]][[6]])
}

#model1
    df_models1_bind <- df_models1_bind[c(-1)]
    df_models1_bind <- df_models1_bind %>% dplyr::rename(A = age, predrate = rate)
    df_models1_bind <- df_models1_bind[,c(10,1,2,3,4,5,6,7,8,9,11)]
    predrates_tscv_bind_m1 <- merge(newlist_tscv5[[i]][[21]][["rates_tscv"]], df_models1_bind, all = TRUE)
    #predrates_tscv_bind_m1 <- left_join(df_models1_bind, newlist_tscv5[[i]][[21]][["rates_tscv"]], by = c("P", "A")) 
    predrates_tscv_bind_m1$RMSE <- ((predrates_tscv_bind_m1$rate - predrates_tscv_bind_m1$predrate)^2)
    
    RMSE_year_m2 <- predrates_tscv_bind_m1 %>% group_by(P) %>% dplyr::summarise(RMSE_year = sqrt(mean(RMSE)))
    predrates_tscv_RMSE_m1 <- sqrt(mean(RMSE_year_m2$RMSE_year^2, na.rm = TRUE))
    predrates_tscv_RMSE_m1 <- as.data.frame(predrates_tscv_RMSE_m1)
    colnames(predrates_tscv_RMSE_m1) <- c("RMSE")
    predrates_tscv_RMSE_m1$modelno <- rep(1, times = nrow(predrates_tscv_RMSE_m1))

#model2
df_models2_bind <- df_models2_bind[c(-1)]
    df_models2_bind <- df_models2_bind %>% dplyr::rename(A = age, predrate = rate)
    df_models2_bind <- df_models2_bind[,c(10,1,2,3,4,5,6,7,8,9,11)]
    predrates_tscv_bind_m2 <- merge(newlist_tscv5[[i]][[21]][["rates_tscv"]], df_models2_bind, all = TRUE)
    #predrates_tscv_bind_m2 <- left_join(df_models2_bind, newlist_tscv5[[i]][[21]][["rates_tscv"]], by = c("P", "A"))
    predrates_tscv_bind_m2$RMSE <- ((predrates_tscv_bind_m2$rate - predrates_tscv_bind_m2$predrate)^2)
    
    RMSE_year_m2 <- predrates_tscv_bind_m2 %>% group_by(P) %>% dplyr::summarise(RMSE_year = sqrt(mean(RMSE)))
    predrates_tscv_RMSE_m2 <- sqrt(mean(RMSE_year_m2$RMSE_year^2, na.rm = TRUE))
    predrates_tscv_RMSE_m2 <- as.data.frame(predrates_tscv_RMSE_m2)
    colnames(predrates_tscv_RMSE_m2) <- c("RMSE")
     predrates_tscv_RMSE_m2$modelno <- rep(2, times = nrow(predrates_tscv_RMSE_m2))
    
#model3
 df_models3_bind <- df_models3_bind[c(-1)]
    df_models3_bind <- df_models3_bind %>% dplyr::rename(A = age, predrate = rate)
    df_models3_bind <- df_models3_bind[,c(10,1,2,3,4,5,6,7,8,9,11)]
    predrates_tscv_bind_m3 <- merge(newlist_tscv5[[i]][[21]][["rates_tscv"]], df_models3_bind, all = TRUE)
    #predrates_tscv_bind_m3 <- left_join(df_models3_bind, newlist_tscv5[[i]][[21]][["rates_tscv"]], by = c("P", "A"))
    predrates_tscv_bind_m3$RMSE <- ((predrates_tscv_bind_m3$rate - predrates_tscv_bind_m3$predrate)^2)
    
    RMSE_year_m3 <- predrates_tscv_bind_m3 %>% group_by(P) %>% dplyr::summarise(RMSE_year = sqrt(mean(RMSE)))
    predrates_tscv_RMSE_m3 <- sqrt(mean(RMSE_year_m3$RMSE_year^2, na.rm = TRUE))
    predrates_tscv_RMSE_m3 <- as.data.frame(predrates_tscv_RMSE_m3)
    colnames(predrates_tscv_RMSE_m3) <- c("RMSE")
    predrates_tscv_RMSE_m3$modelno <- rep(3, times = nrow(predrates_tscv_RMSE_m3))
    
#model4
 df_models4_bind <- df_models4_bind[c(-1)]
    df_models4_bind <- df_models4_bind %>% dplyr::rename(A = age, predrate = rate)
    df_models4_bind <- df_models4_bind[,c(10,1,2,3,4,5,6,7,8,9,11)]
    predrates_tscv_bind_m4 <- merge(newlist_tscv5[[i]][[21]][["rates_tscv"]], df_models4_bind, all = TRUE)
    #predrates_tscv_bind_m4 <- left_join(df_models4_bind, newlist_tscv5[[i]][[21]][["rates_tscv"]], by = c("P", "A")) 
    predrates_tscv_bind_m4$RMSE <- ((predrates_tscv_bind_m4$rate - predrates_tscv_bind_m4$predrate)^2)
    
    RMSE_year_m4 <- predrates_tscv_bind_m4 %>% group_by(P) %>% dplyr::summarise(RMSE_year = sqrt(mean(RMSE)))
    predrates_tscv_RMSE_m4 <- sqrt(mean(RMSE_year_m4$RMSE_year^2, na.rm = TRUE))
    predrates_tscv_RMSE_m4 <- as.data.frame(predrates_tscv_RMSE_m4)
    colnames(predrates_tscv_RMSE_m4) <- c("RMSE")
    predrates_tscv_RMSE_m4$modelno <- rep(4, times = nrow(predrates_tscv_RMSE_m4))
    
#model5
df_models5_bind <- df_models5_bind[c(-1)]
    df_models5_bind <- df_models5_bind %>% dplyr::rename(A = age, predrate = rate)
    df_models5_bind <- df_models5_bind[,c(10,1,2,3,4,5,6,7,8,9,11)]
    predrates_tscv_bind_m5 <- merge(newlist_tscv5[[i]][[21]][["rates_tscv"]], df_models5_bind, all = TRUE)
    #predrates_tscv_bind_m5 <- left_join(df_models5_bind, newlist_tscv5[[i]][[21]][["rates_tscv"]], by = c("P", "A")) 
    predrates_tscv_bind_m5$RMSE <- ((predrates_tscv_bind_m5$rate - predrates_tscv_bind_m5$predrate)^2)
    
    RMSE_year_m5 <- predrates_tscv_bind_m5 %>% group_by(P) %>% dplyr::summarise(RMSE_year = sqrt(mean(RMSE)))
    predrates_tscv_RMSE_m5 <- sqrt(mean(RMSE_year_m5$RMSE_year^2, na.rm = TRUE))
    predrates_tscv_RMSE_m5 <- as.data.frame(predrates_tscv_RMSE_m5)
    colnames(predrates_tscv_RMSE_m5) <- c("RMSE")    
    predrates_tscv_RMSE_m5$modelno <- rep(5, times = nrow(predrates_tscv_RMSE_m5))
    
#model6   
df_models6_bind <- df_models6_bind[c(-1)]
    df_models6_bind <- df_models6_bind %>% dplyr::rename(A = age, predrate = rate)
    df_models6_bind <- df_models6_bind[,c(10,1,2,3,4,5,6,7,8,9,11)]
    predrates_tscv_bind_m6 <- merge(newlist_tscv5[[i]][[21]][["rates_tscv"]], df_models6_bind, all = TRUE)
    #predrates_tscv_bind_m6 <- left_join(df_models6_bind, newlist_tscv5[[i]][[21]][["rates_tscv"]], by = c("P", "A"))
    predrates_tscv_bind_m6$RMSE <- ((predrates_tscv_bind_m6$rate - predrates_tscv_bind_m6$predrate)^2)
    
    RMSE_year_m6 <- predrates_tscv_bind_m6 %>% group_by(P) %>% dplyr::summarise(RMSE_year = sqrt(mean(RMSE)))
    predrates_tscv_RMSE_m6 <- sqrt(mean(RMSE_year_m6$RMSE_year^2, na.rm = TRUE))
    predrates_tscv_RMSE_m6 <- as.data.frame(predrates_tscv_RMSE_m6)
    colnames(predrates_tscv_RMSE_m6) <- c("RMSE")
    predrates_tscv_RMSE_m6$modelno <- rep(6, times = nrow(predrates_tscv_RMSE_m6))

RMSE_bind_tscv5 <- rbind(predrates_tscv_RMSE_m1, predrates_tscv_RMSE_m2, predrates_tscv_RMSE_m3, predrates_tscv_RMSE_m4, predrates_tscv_RMSE_m5, predrates_tscv_RMSE_m6)

#RMSE_bind_tscv5 <- rbind(predrates_tscv_bind_m1, predrates_tscv_bind_m2, predrates_tscv_bind_m3, predrates_tscv_bind_m4, predrates_tscv_bind_m5, predrates_tscv_bind_m6) #use to check what dfs look like prior to overall RMSE calc

RMSEresults_tscv5[[i]] <-  RMSE_bind_tscv5

}
names(RMSEresults_tscv5) <- names(newlist_tscv5)

```

#Run all 6 GAMS and produce plots for visual inspection prior to selecting ones that need manual selection
```{r}
allGAMfits <- vector(mode = "list", length(res.list))

for(i in 1:length(res.list)){

models_all <- gammodels_all_f(res.list[[i]])
models_all$age <- as.factor(models_all$age)
models_all <- models_all[c(-1)]
models_all <- models_all %>% dplyr::rename(A = age)

observedrates <- rates_f(res.list[[i]])
observedrates$A <- as.factor(observedrates$A)

m1 <- filter(models_all, status == "predicted, k=10, penalty=1")
m2 <- filter(models_all, status == "predicted, k=10, penalty=2")
m3 <- filter(models_all, status == "predicted, k=10, penalty=2,1")
m4 <- filter(models_all, status == "predicted, k=5, penalty=1")
m5 <- filter(models_all, status == "predicted, k=5, penalty=2")
m6 <- filter(models_all, status == "predicted, k=5, penalty=2,1")

m1 <- dplyr::bind_rows(observedrates, m1)
m2 <- dplyr::bind_rows(observedrates, m2)
m3 <- dplyr::bind_rows(observedrates, m3)
m4 <- dplyr::bind_rows(observedrates, m4)
m5 <- dplyr::bind_rows(observedrates, m5)
m6 <- dplyr::bind_rows(observedrates, m6)

plot_freey_m1 <- ggplot(m1, aes(x=P, y=rate)) + geom_point(aes(color=status)) + labs(x = "Year of diagnosis", y = "Rate per million person-years", title = paste0("Projected cancer incidence rates for", " ", paste0(names(res.list[i])), ", free y axes and 95% CIs")) + facet_wrap(m1$A, scales = "free_y") + theme(legend.position = "bottom") #+ theme(axis.text.y = element_blank())
plot_CI_freey_m1 <- plot_freey_m1 + geom_ribbon(aes(ymin=lwr_ci, ymax=upr_ci), linetype=2, alpha=0.1)

plot_freey_m2 <- ggplot(m2, aes(x=P, y=rate)) + geom_point(aes(color=status)) + labs(x = "Year of diagnosis", y = "Rate per million person-years", title = paste0("Projected cancer incidence rates for", " ", paste0(names(res.list[i])), ", free y axes and 95% CIs")) + facet_wrap(m2$A, scales = "free_y") + theme(legend.position = "bottom") #+ theme(axis.text.y = element_blank())
plot_CI_freey_m2 <- plot_freey_m2 + geom_ribbon(aes(ymin=lwr_ci, ymax=upr_ci), linetype=2, alpha=0.1)

plot_freey_m3 <- ggplot(m3, aes(x=P, y=rate)) + geom_point(aes(color=status)) + labs(x = "Year of diagnosis", y = "Rate per million person-years", title = paste0("Projected cancer incidence rates for", " ", paste0(names(res.list[i])), ", free y axes and 95% CIs")) + facet_wrap(m3$A, scales = "free_y") + theme(legend.position = "bottom") #+ theme(axis.text.y = element_blank())
plot_CI_freey_m3 <- plot_freey_m3 + geom_ribbon(aes(ymin=lwr_ci, ymax=upr_ci), linetype=2, alpha=0.1)

plot_freey_m4 <- ggplot(m4, aes(x=P, y=rate)) + geom_point(aes(color=status)) + labs(x = "Year of diagnosis", y = "Rate per million person-years", title = paste0("Projected cancer incidence rates for", " ", paste0(names(res.list[i])), ", free y axes and 95% CIs")) + facet_wrap(m4$A, scales = "free_y") + theme(legend.position = "bottom") #+ theme(axis.text.y = element_blank())
plot_CI_freey_m4 <- plot_freey_m4 + geom_ribbon(aes(ymin=lwr_ci, ymax=upr_ci), linetype=2, alpha=0.1)

plot_freey_m5 <- ggplot(m5, aes(x=P, y=rate)) + geom_point(aes(color=status)) + labs(x = "Year of diagnosis", y = "Rate per million person-years", title = paste0("Projected cancer incidence rates for", " ", paste0(names(res.list[i])), ", free y axes and 95% CIs")) + facet_wrap(m5$A, scales = "free_y") + theme(legend.position = "bottom") #+ theme(axis.text.y = element_blank())
plot_CI_freey_m5 <- plot_freey_m5 + geom_ribbon(aes(ymin=lwr_ci, ymax=upr_ci), linetype=2, alpha=0.1)

plot_freey_m6 <- ggplot(m6, aes(x=P, y=rate)) + geom_point(aes(color=status)) + labs(x = "Year of diagnosis", y = "Rate per million person-years", title = paste0("Projected cancer incidence rates for", " ", paste0(names(res.list[i])), ", free y axes and 95% CIs")) + facet_wrap(m6$A, scales = "free_y") + theme(legend.position = "bottom") #+ theme(axis.text.y = element_blank())
plot_CI_freey_m6 <- plot_freey_m6 + geom_ribbon(aes(ymin=lwr_ci, ymax=upr_ci), linetype=2, alpha=0.1)

allGAMfits[[i]] <- list(plot_CI_freey_m1, plot_CI_freey_m2, plot_CI_freey_m3, plot_CI_freey_m4, plot_CI_freey_m5, plot_CI_freey_m6)

names(allGAMfits) <- names(res.list)
}

pdf("REDACTED", width = 11, height = 8)
for (i in 1:168) { #168 for 28 groups, printing 6 plots each
  print(allGAMfits[[i]])
}
dev.off()
```


#5. Create table with lowest RMSE value/model fit for each of the 28 cancertype/sex combinations. Allow this to be pulled and referenced so that right m1 can be run within final prediction and plotting in #6
```{r}
#Combine all RMSE results
RMSEresults_bind_tscv5 <- ldply(RMSEresults_tscv5, data.frame)
RMSEresults_bind_tscv5 <- RMSEresults_bind_tscv5 %>% dplyr::rename(group = .id)
RMSEresults_bind_tscv5$id <- rep(1:28, each = 6, times = 1)

#Create table with convergence results
convergence_bind <- newlist_tscv5[[1]][[21]][["df_models_tscv5"]][[7]]

for(i in 2:28){
  convergence_bind <- rbind(convergence_bind, newlist_tscv5[[i]][[21]][["df_models_tscv5"]][[7]])
}
convergence_bind$group <- paste0(RMSEresults_bind_tscv5$group)
convergence_bind$id <- rep(1:28, each = 6, times = 1)

RMSEresults_bind_tscv5 <- left_join(RMSEresults_bind_tscv5, convergence_bind, by = c("group", "modelno", "id"))

write.csv(RMSEresults_bind_tscv5, "RMSEresults_allmodels.csv") #prior to manual selection**

#######
#filter to find min RMSE for each group and save
RMSEresults_min_tscv5 <- RMSEresults_bind_tscv5 %>% group_by(group) %>% slice(which.min(RMSE)) %>% select(!RMSE) %>% mutate(manualselection = rep("No"))

#Inspect pdf of all GAM fits to determine changes. manually choose model for select groups where initially chosen one was not most appropriate - March 2022

#Females
RMSEresults_min_tscv5$modelno[RMSEresults_min_tscv5$group == "ALL_Female"] <- 5 #was model #6 prior
RMSEresults_min_tscv5$manualselection[RMSEresults_min_tscv5$group == "ALL_Female"] <- "Yes"

RMSEresults_min_tscv5$modelno[RMSEresults_min_tscv5$group == "Neuroblastoma_Female"] <- 5 #was model #4 prior
RMSEresults_min_tscv5$manualselection[RMSEresults_min_tscv5$group == "Neuroblastoma_Female"] <- "Yes"

RMSEresults_min_tscv5$modelno[RMSEresults_min_tscv5$group == "Renal tumours_Female"] <- 2 #was model #1 prior
RMSEresults_min_tscv5$manualselection[RMSEresults_min_tscv5$group == "Renal tumours_Female"] <- "Yes"

RMSEresults_min_tscv5$modelno[RMSEresults_min_tscv5$group == "Hepatic tumours_Female"] <- 3 #was model #1 prior
RMSEresults_min_tscv5$manualselection[RMSEresults_min_tscv5$group == "Hepatic tumours_Female"] <- "Yes"

RMSEresults_min_tscv5$modelno[RMSEresults_min_tscv5$group == "Soft tissue sarcomas_Female"] <- 5 #was model #4 prior
RMSEresults_min_tscv5$manualselection[RMSEresults_min_tscv5$group == "Soft tissue sarcomas_Female"] <- "Yes"

RMSEresults_min_tscv5$modelno[RMSEresults_min_tscv5$group == "Other epithelial and unspecified neoplasms_Female"] <- 5 #was model #2 prior
RMSEresults_min_tscv5$manualselection[RMSEresults_min_tscv5$group == "Other epithelial and unspecified neoplasms_Female"] <- "Yes"

#### Male
RMSEresults_min_tscv5$modelno[RMSEresults_min_tscv5$group == "AML and other leukemias_Male"] <- 5 #was model #4 prior
RMSEresults_min_tscv5$manualselection[RMSEresults_min_tscv5$group == "AML and other leukemias_Male"] <- "Yes"

RMSEresults_min_tscv5$modelno[RMSEresults_min_tscv5$group == "Non-Hodgkin lymphomas and other lymphomas_Male"] <- 4 #was model #1 prior
RMSEresults_min_tscv5$manualselection[RMSEresults_min_tscv5$group == "Non-Hodgkin lymphomas and other lymphomas_Male"] <- "Yes"

RMSEresults_min_tscv5$modelno[RMSEresults_min_tscv5$group == "Neuroblastoma_Male"] <- 2 #was model #1 prior
RMSEresults_min_tscv5$manualselection[RMSEresults_min_tscv5$group == "Neuroblastoma_Male"] <- "Yes"

RMSEresults_min_tscv5$modelno[RMSEresults_min_tscv5$group == "Hepatic tumours_Male"] <- 2 #was model #3 prior
RMSEresults_min_tscv5$manualselection[RMSEresults_min_tscv5$group == "Hepatic tumours_Male"] <- "Yes"

RMSEresults_min_tscv5$modelno[RMSEresults_min_tscv5$group == "Bone tumours_Male"] <- 6 #was model #3 prior
RMSEresults_min_tscv5$manualselection[RMSEresults_min_tscv5$group == "Bone tumours_Male"] <- "Yes"

RMSEresults_min_tscv5$modelno[RMSEresults_min_tscv5$group == "Germ cell tumours_Male"] <- 5 #was model #6 prior
RMSEresults_min_tscv5$manualselection[RMSEresults_min_tscv5$group == "Germ cell tumours_Male"] <- "Yes"

RMSEresults_min_tscv5$modelno[RMSEresults_min_tscv5$group == "Other epithelial and unspecified neoplasms_Male"] <- 5 #was model #4 prior
RMSEresults_min_tscv5$manualselection[RMSEresults_min_tscv5$group == "Other epithelial and unspecified neoplasms_Male"] <- "Yes"

RMSEresults_min_tscv5 <- left_join(RMSEresults_min_tscv5, RMSEresults_bind_tscv5, by = c("group", "modelno", "id", "converge"))

RMSEresults_min_tscv5 <- RMSEresults_min_tscv5[order(RMSEresults_min_tscv5$id),]
write.csv(RMSEresults_min_tscv5, "RMSEresults_selectedmodels.csv")

```


#######################################################################################
#### Predict future incidence rates and plot #####
#######################################################################################

#6. For the CHOSEN gam model fit chosen from tscv - predict for each cancer type/sex incidence count frame and plot
```{r}

plotrates_list <- vector(mode = "list", length(res.list)) #res.list contains the original 28 incidence count dataframes for all cancertype/sex groupings

for(i in 1:length(res.list)){

observedrates <- rates_f(res.list[[i]])
m1 <- gammodel_f(data = res.list[[i]], modelno = RMSEresults_min_tscv5$modelno[i]) #using model selection from tscv predicting 5th year ahead
#predicts 2020-2040 only:
#Pfr_export <- Pfr_export_f(res.list[[i]]) 
#predrates <- predrates_f(res.list[[i]])
#plotrates <- plotrates_f(res.list[[i]])
#plot <- plot_f(res.list[[i]]) #plot with predicted rates
#plot_freey <- plot_freey_f(res.list[[i]]) #plot with predicted rates, free y axis for each age
#plot_CI <- plot_CI_f(res.list[[i]]) #plot with predicted rates + cipred2 95% CI
#plot_CI_freey <- plot_CI_freey_f(res.list[[i]]) #plot with predicted rates, free y axis for each age + cipred2 95% CI

#predict back and forward (1950-2040)
Pfr_export_all <- Pfr_export_all_f(res.list[[i]])
predrates_all <- predrates_all_f(res.list[[i]])
plotrates_all <- plotrates_all_f(res.list[[i]])
# 
# plot_all <- plot_all_f(res.list[[i]])
# plot_all_freey <- plot_all_freey_f(res.list[[i]])
# plot_CI_all <- plot_CI_all_f(res.list[[i]])
# plot_CI_all_freey <- plot_CI_all_freey_f(res.list[[i]])

#plotrates_list[[i]] <- list(plot_all = plot_all, plot_all_freey = plot_all_freey, plot_CI_all = plot_CI_all, plot_CI_all_freey = plot_CI_all_freey) #for exporting plots of whole observed period+predicted period

#plotrates_list[[i]] <- list(predrates_all = predrates_all) #for exporting predicted rates csv

plotrates_list[[i]] <- list(plotrates_all = plotrates_all) #for plotsums graphs
}
names(plotrates_list) <- names(res.list)

```

#export plots
```{r, include = FALSE}

pdf("REDACTED", width = 11, height = 8)
for (i in 1:112) { #112 for 28 groups, printing 4 plots each
  print(plotrates_list[[i]])
}
dev.off()
```

#export csv of predicted rates
```{r}
#run plotrates_list with just dataframes first (predrates_all)
predrates_all_bind <- ldply(plotrates_list, data.frame)
colnames(predrates_all_bind) <- c("group", "P", "lograte", "logSE", "A", "rate", "SE", "lower", "upper", "status")

write.csv(predrates_all_bind, "GAM_incidence_results.csv")

```

