---
title: "MSM_all models_70to89"
output: html_document
---

======================================================================

# 2018 Institute for Clinical Evaluative Sciences. All rights reserved.

# TERMS OF USE:

##Not for distribution.## This code and data is provided to the user solely for its own non-commerical use by individuals and/or not-for-profit corporations. Users shall not distribute without express written permission from the Institute for Clinical Evaluative Sciences.

##Not-for-profit.## This code and data may not be used in connection with profit generating activities.

##No liability.## The Institute for Clinical Evaluative Sciences makes no warranty or representation regarding the fitness, quality or reliability of this code and data.

##No Support.## The Institute for Clinical Evaluative Sciences will not provide any technological, educational or informational support in connection with the use of this code and data.

##Warning.## By receiving this code and data, user accepts these terms, and uses the code and data, solely at its own risk.

======================================================================

---

This file runs multi-state parametric survival models to examine survival after childhood cancer diagnosis.

This coding process was applied to two separate cohorts: 1) children diagnosed with cancer between *1970-1989*, captured in the Ontario Cancer Registry (data held at ICES); and 2) children diagnosed with cancer between *1990-2019*, captured in the Pediatric Oncology Group of Ontario Networked Information System (data held at ICES).

One line (line 104) requires manual selection, depending on the years of data available for analysis.

# 01 Load packages

```{r setup, include=FALSE}
rm(list=ls())

library(dplyr)
library(flexsurv)
library(flexsurvcure)
library(survminer)
library(ggplot2)
library(msm)
library(mstate)
library(reshape2)
library(knitr)
library(abind)
library(data.table)
library(openxlsx)
library(gridExtra)
library(tidyr)
library(lubridate)
library(plyr)

source('functions_msm.R')

output_file_directory <- "REDACTED"

```

#Assign state info
```{r}
v_names_states1 <- c("Diagnosis", "CRE", "Death")
n_states        <- length(v_names_states1)                        # number of states 

n_t_yr   <- 101                                  # time horizon in years 
c_l_yr   <- 1                                    # cycle length (annual)
times_yr <- seq(from = 0, to = n_t_yr, by = c_l_yr) # sequence of cycle days to be considered in the model (annual)

covariates <- c("sex", "age", "dxyear", "ICCC_regroup")

v_n <- c("Diagnosis", "CRE", "Death")
v_s <- c("Diagnosis_s", "CRE_s", "Death_s")

m_t <- transMat(x = list(c(2,3),
                         c(3),
                         c()),
                names = v_n)


# Transition matrix

m_P_diag <- matrix(NA, n_states, n_states, dimnames = list(v_names_states1, v_names_states1))
m_P_diag["Diagnosis", "CRE"] <- 1
m_P_diag["Diagnosis", "Death"] <- 2
m_P_diag["CRE", "Death"] <- 3
#layout.fig <- c(2, 1)
#plotmat(t(m_P_diag), t(layout.fig), self.cex = 0.5, curve = 0, arr.pos = 0.76,  
        #latex = T, arr.type = "curved", relsize = 0.85, box.prop = 0.8, 
        #cex = 0.1, box.cex = 0.7, lwd = 1)
```


```{r}
MSMdata <- read.csv("REDACTED")

timedata <- MSMdata %>% select(Progression_time, Death_time)
timedata$Progressionfree <- rep(0, times = nrow(timedata))
timedata <- timedata %>% select(Progressionfree, Progression_time, Death_time) %>% dplyr::rename(Diagnosis = Progressionfree, CRE = Progression_time, Death = Death_time)

status <- MSMdata %>% select(Progressionfree, Progression, Death) %>% dplyr::rename(Diagnosis_s = Progressionfree, CRE_s = Progression, Death_s = Death)

MSMdata <- MSMdata %>% dplyr::rename(Diagnosis_s = Progressionfree, CRE_s = Progression, Death_s = Death, CRE = Progression_time, Death = Death_time, id = ID)

MSMdata$dxyear <- MSMdata$dxyear - 1969 #This line was run when analyzing survival for patients diagnosed between 1970-1989 (Subtracting by 1 to have dxyear of 1970 == 1, 1971 == 2, ...). When analyzing survival for patients diagnosed between 1990-2019, MSMdata$dxyear - 1989 was applied.

#Multistate Modeling (MSM)
## Read in MSM datasets

data_long_covar  <- msprep(time = timedata, status = status, trans = m_t, data = MSMdata, keep = covariates)

v_cov <- c("sex", "age", "dxyear") #covariates assessed in survival models. age and dxyear were continuous variables.
v_cov <- paste(v_cov, collapse = "+")
v_cov <- noquote(v_cov)

models <- c("Exponential", "Weibull", "Gamma", "LogNormal", "Gompertz", "LogLogistic", "Spline", "Exponential.c", "Weibull.c", "Gamma.c", "LogNormal.c", "Gompertz.c", "LogLogistic.c") #distributions to assess (.c is a mixture cure model)

cov.lists <- c(rep(v_cov, 1),
               rep(v_cov, 1),
               rep(v_cov, 1))


#create and store subsetted dataframes for MSM by cancer type
ICCCgroups <- MSMdata %>% dplyr::count(ICCC_regroup)

MSMdfs <- vector(mode = "list")
for(i in 1:dim(ICCCgroups)[1]){
  
df <- filter(data_long_covar, ICCC_regroup == ICCCgroups$ICCC_regroup[i])

MSMdfs[[i]] <- df

names(MSMdfs)[i] <- paste0(ICCCgroups$ICCC_regroup[i])

MSMdfs[[i]] <- as.data.frame(MSMdfs[[i]])

}

```


```{r}
model.list.results <- vector(mode = "list")

for(i in 1:length(MSMdfs)){  

#run all models for all transitions, store in model.list
model.list <- fun.model.list(v_n = v_n,
                             m_t = m_t,
                             df = MSMdfs[[i]],
                             cov.lists = cov.lists)

fit.list <- fun.fit.list(m_t = m_t,
                         model.list = model.list,
                         models = models)

model.list.results[[i]] <- list(model.list = model.list, fit.list = fit.list)

}
names(model.list.results) <- names(MSMdfs)

```


```{r}

#Diagnosis to CRE transition
fit_Dx_CRE_iccc <- vector(mode = "list")

for(i in 1:length(MSMdfs)){

fit_Dx_CRE_iccc_model <- state.model(data = subset(MSMdfs[[i]], to == "2" & from == "1"),
                        covariates = v_cov,
                        title = paste0("Diagnosis to CRE", ", ", names(MSMdfs)[i]))

fit_Dx_CRE_iccc_model_export <- output.f(fit_Dx_CRE_iccc_model)
fit_Dx_CRE_iccc_model_export_df <- ldply(fit_Dx_CRE_iccc_model_export, data.frame)
fit_Dx_CRE_iccc[[i]] <- fit_Dx_CRE_iccc_model_export_df

}
names(fit_Dx_CRE_iccc) <- names(MSMdfs)

#Diagnosis to Death transition
fit_Dx_D_iccc <- vector(mode = "list")

for(i in 1:length(MSMdfs)){  

fit_Dx_D_iccc_model <- state.model(data = subset(MSMdfs[[i]], to == "3" & from == "1"),
                        covariates = v_cov,
                        title = paste0("Diagnosis to Death", ", ", names(MSMdfs)[i]))

fit_Dx_D_iccc_model_export <- output.f(fit_Dx_D_iccc_model)
fit_Dx_D_iccc_model_export_df <- ldply(fit_Dx_D_iccc_model_export, data.frame)
fit_Dx_D_iccc[[i]] <- fit_Dx_D_iccc_model_export_df
}
names(fit_Dx_D_iccc) <- names(MSMdfs)


#CRE to Death transition
fit_CRE_D_iccc <- vector(mode = "list")

for(i in 1:length(MSMdfs)){

fit_CRE_D_iccc_model <- state.model(data = subset(MSMdfs[[i]], to == "3" & from == "2"),
                        covariates = v_cov,
                        title = paste0("CRE to Death", ", ", names(MSMdfs)[i]))

fit_CRE_D_iccc_model_export <- output.f(fit_CRE_D_iccc_model)
fit_CRE_D_iccc_model_export_df <- ldply(fit_CRE_D_iccc_model_export, data.frame)

fit_CRE_D_iccc[[i]] <- fit_CRE_D_iccc_model_export_df
}
names(fit_CRE_D_iccc) <- names(MSMdfs)

```

## Extract model information and save as csv files
### Save AIC and BIC for all models

```{r}
# Create a vector for all transition names   
v_transition_names <- c("model_Diagnosis_CRE", "model_Diagnosis_Death", "model_CRE_Death")

#### run models.AICBIC function for each cancer type
df_all_AICBIC.results <- vector(mode = "list")

for(i in 1:length(model.list.results)){  

df_all_AICBIC <- models.AICBIC(models, v_transition_names, model.list.results[[i]][["fit.list"]])

df_all_AICBIC.results[[i]] <- list(df_all_AICBIC = df_all_AICBIC)

}
names(df_all_AICBIC.results) <- names(model.list.results)

# write AIC and BIC for all models to a csv
df_all_AICBIC_combined <- ldply(df_all_AICBIC.results, data.frame)
colnames(df_all_AICBIC_combined) <- c("ICCC_regroup", "transition", "model", "AIC", "BIC", "cov") 

write.csv(df_all_AICBIC_combined, "df_all_AICBIC_combined.csv")

```

#Combine all dfAICBIC results to create summary file for export
#Choose model with best estimated predictive ability from AIC/BIC values
#Model must have converged and have a covariance matrix (cov == 1)
```{r}
df_all_models_best_aic <- df_all_AICBIC_combined %>% 
  dplyr::select(-BIC) %>% 
  filter(cov == 1) %>%  # remove rows with no cov matrix 
  filter(!model %in% c("Exponential.c", "Weibull.c" , "Gamma.c","LogNormal.c","Gompertz.c", "LogLogistic.c")) %>% #remove results from mixture cure models
  group_by(ICCC_regroup, transition) %>% 
  na.omit(AIC) %>% 
  dplyr::summarise(min_AIC   = min(AIC), model_AIC = model[AIC == min(AIC)])

df_all_models_best_bic <- df_all_AICBIC_combined %>% 
  dplyr::select(-AIC) %>%
  filter(cov == 1) %>%                     # remove rows with no cov matrix
  filter(!model %in% c("Exponential.c", "Weibull.c" , "Gamma.c","LogNormal.c","Gompertz.c", "LogLogistic.c")) %>% #remove results from mixture cure models
  na.omit(BIC) %>% 
  group_by(ICCC_regroup, transition) %>% 
  dplyr::summarise(min_BIC   = min(BIC),
            model_BIC = model[BIC == min(BIC)])

df_all_models_best <- left_join(df_all_models_best_aic, df_all_models_best_bic) %>% 
  mutate(same_model = ifelse(model_AIC == model_BIC, "TRUE", "FALSE"))

#Write best AIC and BIC models for each transition to a csv

write.csv(df_all_models_best, "df_all_models_best.csv")
```

###Extract estimates and covariance matrices from all models - by cancer type

```{r}
extract.results <- vector(mode = "list")

for(h in 1:length(MSMdfs)){  

model.list <- fun.model.list(v_n = v_n, #run model.list again
                            m_t = m_t,
                             df = MSMdfs[[h]],
                             cov.lists = cov.lists)
names(model.list) <- v_transition_names

# Data frame with all the available models per transition
df_models_by_transition <- df_all_AICBIC_combined %>% 
  mutate(no_model = ifelse(!is.na(AIC) | !is.na(BIC), 1, 0)) %>%    # if AIC OR BIC isn't NA, no_model = 1
  filter(no_model == 1 & ICCC_regroup == names(MSMdfs)[h]) %>%     # filter models that have AIC and/or BIC values
  dplyr::select(transition, model)                                 # for models with AIC and BIC values, 

l_model_by_transition <- vector(mode = "list", length = length(v_transition_names)) # to store available models for each transition
names(l_model_by_transition) <- v_transition_names                                  # name each transition

df_model_knots <- data.frame(transition = as.character(),
                             knots = as.numeric())

# Fill l_model_by_transition with loop to extract the available models for each transition
for (j in 1:length(v_transition_names)) {              # for each transition
  df_available_models <- df_models_by_transition %>%   
    filter(transition == v_transition_names[j])        # extract available models  
  
  l_model_by_transition[[j]] <- df_available_models    # store in list 
}

# Creating 2 empty lists to store estimates and cov matrix 
l_all_models_by_transition_est <- vector(mode   = "list", 
                                         length = length(v_transition_names))

l_all_models_by_transition_cov <- vector(mode   = "list",
                                         length = length(v_transition_names))

l_all_models_by_transition_knots <- vector(mode   = "list",
                                         length = length(v_transition_names))
# Name each element by transition
names(l_all_models_by_transition_knots) <- names(l_all_models_by_transition_est) <- names(l_all_models_by_transition_cov) <- v_transition_names

# Open transition loop
for (transition in 1:length(model.list)){ 
  
  transition_name    <- names(model.list)[transition]             # Name of the transition
  l_transition       <- model.list[[transition_name]]             # List of all models for transition 
  v_available_models <- l_model_by_transition[[transition]]$model # Vector of available models for transition
  
  df_all_models_by_transition_est <- c()                          # empty dataframe to catch model estimates
  l_all_models_cov <- vector(mode   = "list",                     # empty list to catch model variance covariance matrices
                             length = length(v_available_models))
  
  names(l_all_models_cov) <- v_available_models                   # Name elements of list 
  
  model_knots <- data.frame(transition = transition_name,
                            knots = model.list[[transition]]$knots)
  names(model_knots) <- c("transition", "knots")
  df_model_knots <- bind_rows(df_model_knots, model_knots)
  
  # Open available model loop (for each transition)
  for (available_model in 1:ifelse(length(v_available_models)==0, 1, length(v_available_models))){ #ifelse catches if no available models
    
    available_model_name <- v_available_models[available_model]      # Name of the model
    l_model              <- l_transition[[available_model_name]]     # Stores model
    m_model_est          <- l_model[["res"]]                         # Results of the model
    m_model_cov          <- l_model[["cov"]]                         # Var-Cov matrix of the model
    
    v_cov_names <- names(m_model_est[,1])                            # Names of parameters and covariates
    if (is.null(v_cov_names)){v_cov_names <- NA}                     # if empty, NA
    
    m_model_est <- m_model_est[,1]                                   # Stores parameters and coefficients
    if (is.null(m_model_est)){m_model_est <- NA}                     # if empty, NA
    
    # Get SE 
    SE_trans <- c()
    for (i in 1:length(m_model_est)){
      SE_trans[i]<- ifelse(tryCatch({is.null(m_model_cov[i, i])}, 
                      error = function(e){NA}), 
                      NA,                        # if error (ex. incorrect # of dimensions) take NA 
                      sqrt(m_model_cov[i, i]))   # otherwise take sqrt of value
      
    }
    
    # Calculate p-value
    pvalue <- round(2*pnorm(-abs(m_model_est/SE_trans)), 3)
    
    # Create a data frame with the estimates of the model
    df_model_est <- data.frame(transition   = transition_name,       # transition name
                               model        = available_model_name,  # model name
                               covariate    = v_cov_names,           # parameter and covariate names
                               estimate     = m_model_est,           # parameter and covariate coefficient values
                               SE           = SE_trans,              # SE
                               pvalue       = pvalue,                # pvalue
                               row.names    = NULL)                  # no row names
    
    df_all_models_by_transition_est   <- bind_rows(df_all_models_by_transition_est, df_model_est) # combines estimates for all models 
    l_all_models_cov[available_model] <- list(m_model_cov)                                        # combines covariance matrices for all models 
  } # end model loop
  
  l_all_models_by_transition_est[transition] <- list(df_all_models_by_transition_est)   # store estimates for all transitions
  l_all_models_by_transition_cov[transition] <- list(l_all_models_cov)                  # store cov matrices for all transitions
  l_all_models_by_transition_knots[transition] <- list(df_model_knots)
  
} # end transition loop

 #### Extract estimates and covariance matrices from selected models
 
# Data frame with the best model of each transition defined by AIC
df_selected_models_best <- df_all_models_best %>%
  filter(ICCC_regroup == names(MSMdfs)[h]) %>%
  dplyr::select(transition, model = model_AIC)

# Empty lists to hold selected models est and covs
l_selected_models_by_transition_est <- vector(mode   = "list", 
                                              length = length(v_transition_names))
l_selected_models_by_transition_cov <- vector(mode   = "list", 
                                              length = length(v_transition_names))

# Name each element of the lists with the transition names
names(l_selected_models_by_transition_est) <- v_transition_names
names(l_selected_models_by_transition_cov) <- v_transition_names

for (transition in 1:length(v_transition_names)) { 
  
  #Name of the transition
  transition_name <- names(l_all_models_by_transition_est)[transition]
  
  #Name of the best model of the transition
  best_model <- df_selected_models_best %>%
    filter(transition == transition_name)
  
  #Data frame with the best model of the transition
  tryCatch({df_selected_models_by_transition <- l_all_models_by_transition_est[[transition]] %>%
    filter(model %in% best_model$model)
  },
  error = function(e){
    print(e$message)
    return(df_selected_models_by_transition)}
  )

  l_all_models_cov_matrix <- l_all_models_by_transition_cov[[transition_name]]
  
  l_selected_models_cov_matrix <- l_all_models_cov_matrix[best_model$model]

  
  #List with a data frame by each transition with the best model
  l_selected_models_by_transition_est[transition] <- list(df_selected_models_by_transition)
  
  l_selected_models_by_transition_cov[transition] <- list(l_selected_models_cov_matrix)
  
}

extract.results[[h]] <- list(l_all_models_by_transition_est = l_all_models_by_transition_est, 
                             l_all_models_by_transition_knots = l_all_models_by_transition_knots, 
                             l_selected_models_by_transition_est = l_selected_models_by_transition_est, 
                             l_all_models_by_transition_cov = l_all_models_by_transition_cov, 
                             l_selected_models_by_transition_cov = l_selected_models_by_transition_cov)

}
names(extract.results) <- names(MSMdfs)

```

#export estimates for all models (i.e., all distributions for all cancer types/transitions)
```{r}
l_all_models_by_transition_est_bind <- vector(mode = "list")

for(i in 1:length(extract.results)){

l_all_models_by_transition_est_binded <- ldply(extract.results[[i]][["l_all_models_by_transition_est"]], data.frame)
l_all_models_by_transition_est_binded$ICCC_regroup <- rep(names(extract.results)[i])

l_all_models_by_transition_est_bind[[i]] <-  l_all_models_by_transition_est_binded

}
names(l_all_models_by_transition_est_bind) <- names(extract.results)

l_all_models_by_transition_est <- ldply(l_all_models_by_transition_est_bind, data.frame) #all together in single csv file

write.csv(l_all_models_by_transition_est, "l_all_models_by_transition_est.csv")

```


#export knots for cancer types/transitions in which a spline model was chosen as best fitting distribution
```{r}
l_all_models_by_transition_knots_bind <- vector(mode = "list")

for(i in 1:length(extract.results)){

l_all_models_by_transition_knots_binded <- as.data.frame(extract.results[[i]][["l_all_models_by_transition_knots"]][["model_CRE_Death"]])
l_all_models_by_transition_knots_binded$ICCC_regroup <- rep(names(extract.results)[i])

l_all_models_by_transition_knots_bind[[i]] <-  l_all_models_by_transition_knots_binded

}
names(l_all_models_by_transition_est_bind) <- names(extract.results)

#rbind all
df_knots_all <- l_all_models_by_transition_knots_bind[[1]]
for(i in 2:14){
  df_knots_all <- rbind(df_knots_all, l_all_models_by_transition_knots_bind[[i]])
}

write.csv(df_knots_all, "df_model_knots.csv")

```


#export estimates from best fitting models (i.e., best fitting distribution for each cancer type/transition)

```{r}

l_selected_models_by_transition_est_bind <- vector(mode = "list")

for(i in 1:length(extract.results)){

l_selected_models_by_transition_est_binded <- ldply(extract.results[[i]][["l_selected_models_by_transition_est"]], data.frame)
l_selected_models_by_transition_est_binded$ICCC_regroup <- rep(names(extract.results)[i])

l_selected_models_by_transition_est_bind[[i]] <-  l_selected_models_by_transition_est_binded

}
names(l_selected_models_by_transition_est_bind) <- names(extract.results)

l_selected_models_by_transition_est <- ldply(l_selected_models_by_transition_est_bind, data.frame) #all together in single csv file

write.csv(l_selected_models_by_transition_est, "l_selected_models_by_transition_est.csv")

```


##export variance covariance matrices for all models (i.e., vcov matrix for all distributions for all cancer types/transitions)
```{r}
for(z in 1:length(extract.results)){ #14 cancer types
  
for (i in 1:length(extract.results[[z]][["l_all_models_by_transition_cov"]])){ #length = 3 for 3 transitions
  
  lapply(1:length(extract.results[[z]][["l_all_models_by_transition_cov"]][[i]]), function(x) 
    
  write.csv(extract.results[[z]][["l_all_models_by_transition_cov"]][[i]][[x]],
                                                                              file = paste0(output_file_directory, '/cov_all/','cov_',
                                                                                            names(extract.results[z]),
                                                                                            ".",
                                                                                            names(extract.results[[z]][["l_all_models_by_transition_cov"]][i]), 
                                                                                            ".", 
                                                                                            names(extract.results[[z]][["l_all_models_by_transition_cov"]][[i]][x]), 
                                                                                            ".csv"),
                                                                              row.names = T))
}
}
```

#export variance covariance matrices from best fitting models (i.e., vcov matrix from best fitting distribution for each cancer type/transition)
```{r}

for(i in 1:length(extract.results)){

# write each cov in l_selected_models_by_transition_cov to a csv file 

  lapply(1:length(extract.results[[i]][["l_selected_models_by_transition_cov"]]),  function (x)  
  
  write.csv(ifelse(length(extract.results[[i]][["l_selected_models_by_transition_cov"]][[x]]) == 1, 
                          extract.results[[i]][["l_selected_models_by_transition_cov"]][[x]][1],    # stops from attaching model name to column names
                          extract.results[[i]][["l_selected_models_by_transition_cov"]][[x]]),      # for empty transitions
            file = paste0(output_file_directory, '/cov_selected/','cov_selected_', names(extract.results[i]), ".", names(extract.results[[i]][["l_selected_models_by_transition_cov"]][x]),'.csv'), # make a csv 
            row.names = T))
}
```

```{r}

```

